---
title: "BBets"
author: "Adrian Cruz"
output: html_document
date: "2024-09-27"
---


This assignment has three questions:  

In a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, month, day.

1. Find the propability that at least two students have the same birthday. If 26 students in the class take the bet how much is Dr.J expected to win/lose. How should Dr.J Wager against the students $1 to make this a fair bet?

The propability of at least two students having the same birthday is 0.81. Dr. Jay is expected to win 16.17 dollars. Dr. J should wager exactly 4.3 dollars against the students to make a fair bet.

2. Assume class size is 26 instead of 35.  With the same bet ($1 ot $1) what is the probability Dr. J wins the bet now.  If 20 of the 26 students make the bet, what is Dr. J's expected profit/loss?

With the same bet, the probability if Dr. J winning is now 0.61. Expected profit for Dr.J if 20 students bet is $4.43.

3. What size class (how many students) makes the $1 againt $1 a fair bet?

The closest amount of students to make the bet fair, $1 to $1, is 23 students with the average profit being close to 0.

## Question 1
```{r setup, include=FALSE}
# Load required libraries
library(ggplot2)
```

```{r simulation, echo=TRUE}
# Function to simulate the birthday problem
simulate_birthday_problem <- function(class_size) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  has_duplicate <- length(birthdays) != length(unique(birthdays))
  return(as.integer(has_duplicate))
}

# Set parameters
num_simulations <- 100
class_size <- 35
results <- numeric(num_simulations)

# Run simulations
for (i in 1:num_simulations) {
  results[i] <- simulate_birthday_problem(class_size)
}

# Calculate the probability of at least one shared birthday
probabilities <- cumsum(results) / (1:num_simulations)
```

```{r histogram, echo=TRUE}
# Plot histogram of probabilities
histogram_plot <- ggplot(data.frame(probabilities), aes(x = probabilities)) +
  geom_histogram(binwidth = 0.05, color = "black", fill = "lightblue") +
  labs(title = "Histogram of Probabilities",
       x = "Probability",
       y = "Frequency") +
  theme_minimal()

print(histogram_plot)
```

```{r summary, echo=TRUE}
# Calculate and print the mean probability
mean_probability <- mean(probabilities)
cat("mean =", mean_probability, "\n")

# Calculate 95% confidence interval
conf_interval <- quantile(probabilities, c(0.025, 0.975))
cat("95% Confidence Interval:", conf_interval, "\n")
```



```{r}
EV= mean_probability*26-(1-mean_probability)*26
print(paste("ev=",EV))
```

```{r}
FAIR=(mean_probability)/(1-mean_probability) 
print(paste("fairbet=",FAIR))
```

The propability of at least two students having the same birthday is 0.81. Dr. Jay is expected to win 16.17 dollars. Dr. J should wager exactly 4.3 dollars against the students to make a fair bet.


## Question 2

```{r}
# Function to simulate the birthday problem
simulate_birthday_problem <- function(class_size= 26) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  has_duplicate <- length(birthdays) != length(unique(birthdays))
  return(as.integer(has_duplicate))
}

# Set parameters
num_simulations <- 100
class_size <- 26
results <- numeric(num_simulations)

# Run simulations
for (i in 1:num_simulations) {
  results[i] <- simulate_birthday_problem(class_size)
}

# Calculate the probability of at least one shared birthday
probabilities <- cumsum(results) / (1:num_simulations)
```

```{r, echo=TRUE}
# Calculate and print the mean probability
mean_probability <- mean(probabilities)
cat("mean =", mean_probability, "\n")
   
# Calculate 95% confidence interval
conf_interval <- quantile(probabilities, c(0.025, 0.975))
cat("95% Confidence Interval:", conf_interval, "\n")
```

```{r}
EV= mean_probability*20-(1-mean_probability)*20
print(paste("ev=",EV))
```

With the same bet, the probability if Dr. J winning is now 0.61. Expected profit for Dr.J if 20 students bet is $4.43.

##Question 3

```{r}
# Function to simulate the birthday problem
simulate_birthday_problem <- function(class_size= 23) {
  birthdays <- sample(1:365, class_size, replace = TRUE)
  has_duplicate <- length(birthdays) != length(unique(birthdays))
  return(as.integer(has_duplicate))
}

# Set parameters
num_simulations <- 100
class_size <- 23
results <- numeric(num_simulations)

# Run simulations
for (i in 1:num_simulations) {
  results[i] <- simulate_birthday_problem(class_size)
}

# Calculate the probability of at least one shared birthday
probabilities <- cumsum(results) / (1:num_simulations)
```

```{r}
# Calculate and print the mean probability
mean_probability <- mean(probabilities)
cat("mean =", mean_probability, "\n")
   
# Calculate 95% confidence interval
conf_interval <- quantile(probabilities, c(0.025, 0.975))
cat("95% Confidence Interval:", conf_interval, "\n")
```

```{r}
EV= mean_probability*20-(1-mean_probability)*20
print(paste("ev=",EV))
```

The closest amount of students to make the bet fair, ($1 to $1), is 23 students with the average profit being close to 0.